The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-09-21 paper2_JNLS/paper.tex
NM

%----------------------------------------------------------------------------%
%% LATEX DEFINITIONS
%----------------------------------------------------------------------------%
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\dx}{\, dx}
% Basic physical parameters and scales
\newcommand{\freqp}{f_p}
\newcommand{\etastd}{\eta_{\text std}}
% Parameters that change crossing the ADC
\newcommand{\depth}{d}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\dupdn}{\depth_{\pm}}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamdn}{\lam_{+}}
\newcommand{\lamupdn}{\lam_{\pm}}
\newcommand{\lamfac}{N}
\newcommand{\drat}{\mathcal{D}}
\newcommand{\dratdn}{\drat_+}
\newcommand{\dratupdn}{\drat_{\pm}}
% Statistical quantities
\newcommand{\En}{\mathcal{E}}
\newcommand{\Mo}{\mathcal{M}}
\newcommand{\skw}{\text{skew}}
\newcommand{\skwdn}{\skw_+}
\newcommand{\var}{\text{var}}
\newcommand{\varup}{\var_-}
\newcommand{\vardn}{\var_+}
% Dimensionless parameters
\newcommand{\ampscale}{\mathcal{A}}
\newcommand{\lengthscale}{\mathcal{L}}
\newcommand{\timescale}{\mathcal{T}}
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}
% Hamiltonian stuff
\newcommand{\uhat}{\hat{u}}
\newcommand{\sympJ}{\mathcal{J}}
\newcommand{\vard}[2]{\frac{\delta #1}{\delta #2}}
\newcommand{\Ham}{\mathcal{H}}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Hupdn}{\Ham^{\pm}}
\newcommand{\Hthree}{\Ham_{3}}
\newcommand{\Htwo}{\Ham_{2}}
%\newcommand{\Fcnl}{\mathcal{F}}
% Truncated stuff
\newcommand{\Proj}{\mathcal{P}_{\Lambda}}
\newcommand{\uL}{u_{\Lambda}}
\newcommand{\HLupdn}{\Ham_{\Lambda}^{\pm}}
\newcommand{\SympL}{\sympJ_{\Lambda}}
% Gibbs and theta
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\Gupdn}{\Gibbs^{\pm}}
\newcommand{\thup}{\theta^{-}}
\newcommand{\thdn}{\theta^{+}}
\newcommand{\thupdn}{\theta^{\pm}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}
\newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
\newcommand{\tavg}[1]{\overline{#1}}

@@

Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain. We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$ corresponding to the physical wavelength $\lambda$ is resolved. In the case $\lamfac = \Lambda$, $\lambda$ corresponds to the smallest resolved wavelength. Often, we will choose $\lamfac = \Lambda$ or $\Lambda/2$.

@@

%----------------------------------------------------------------------------%
% Matching
%----------------------------------------------------------------------------%
\subsection{Matching at the ADC}

Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$. Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e. no nonlinearity and no dispersion). Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
\begin{align}
u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+},
\qquad \mbox{\em Deterministic matching condition}
\end{align}
holding for $x \in [-\pi, \pi]$.
%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.

@@

@@

With the experimental setup described and the theory outlined, we now present results comparing the two. Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table (ref). In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.

>>>>

%----------------------------------------------------------------------------%
%% LATEX DEFINITIONS
%----------------------------------------------------------------------------%
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\dx}{\, dx}
% Basic physical parameters and scales
\newcommand{\freqp}{f_p}
\newcommand{\etastd}{\eta_{\text std}}
% Parameters that change crossing the ADC
\newcommand{\depth}{d}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\dupdn}{\depth_{\pm}}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamdn}{\lam_{+}}
\newcommand{\lamupdn}{\lam_{\pm}}
\newcommand{\lamfac}{N}
\newcommand{\drat}{\mathcal{D}}
\newcommand{\dratdn}{\drat_+}
\newcommand{\dratupdn}{\drat_{\pm}}
% Statistical quantities
\newcommand{\En}{\mathcal{E}}
\newcommand{\Mo}{\mathcal{M}}
\newcommand{\skw}{\text{skew}}
\newcommand{\skwdn}{\skw_+}
\newcommand{\var}{\text{var}}
\newcommand{\varup}{\var_-}
\newcommand{\vardn}{\var_+}
% Dimensionless parameters
\newcommand{\ampscale}{\mathcal{A}}
\newcommand{\lengthscale}{\mathcal{L}}
\newcommand{\timescale}{\mathcal{T}}
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}
% Hamiltonian stuff
\newcommand{\uhat}{\hat{u}}
\newcommand{\sympJ}{\mathcal{J}}
\newcommand{\vard}[2]{\frac{\delta #1}{\delta #2}}
\newcommand{\Ham}{\mathcal{H}}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Hupdn}{\Ham^{\pm}}
\newcommand{\Hthree}{\Ham_{3}}
\newcommand{\Htwo}{\Ham_{2}}
%\newcommand{\Fcnl}{\mathcal{F}}
% Truncated stuff
\newcommand{\Proj}{\mathcal{P}_{\Lambda}}
\newcommand{\uL}{u_{\Lambda}}
\newcommand{\HLupdn}{\Ham_{\Lambda}^{\pm}}
\newcommand{\SympL}{\sympJ_{\Lambda}}
% Gibbs and theta
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\Gupdn}{\Gibbs^{\pm}}
\newcommand{\thup}{\theta^{-}}
\newcommand{\thdn}{\theta^{+}}
\newcommand{\thupdn}{\theta^{\pm}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}
\newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
\newcommand{\tavg}[1]{\overline{#1}}
\newcommand{\transf}{\mathcal{F}}

@@

Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain. We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system. If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength. Often, we will choose an intermediate value such as $\lamfac = \Lambda/2$ or $\Lambda/4$, so that the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.

@@

%----------------------------------------------------------------------------%
% Matching
%----------------------------------------------------------------------------%
\subsection{Matching conditions and the transfer function}

Recall that the abrupt depth change is met by traveling waves at some time $t = T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. The KdV equations govern the evolution and nonlinear dispersion of waves over long timescales, $t \gg \delta^{-2}$. Meanwhile, wave evolution over shorter timescales, $t \ll \delta^{-2}$, is much simpler in that it is simply noninteracting traveling waves with uniform wave speed $c = \sqrt{g \depth}$ (i.e. no nonlinearity and no dispersion). Hence, a short instant after waves pass over the ADC, the waveform does not yet have time to alter significantly, which gives the (deterministic) matching condition
\begin{align}
u(x,t) \vert_{t=0^-} = u(x,t) \vert_{t=0^+},
\qquad \mbox{\em Deterministic matching condition}
\end{align}
holding for $x \in [-\pi, \pi]$. This matching condition is used for the dynamical system \eqref{TKdV}.
%or in Johnson's formulation, it is $\depth^{1/4} \eta$ that matches.

@@

The statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$. In particular, we view $\thup$ as given by the random-state of the incoming wave field. Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by matching statistics at the ADC. In this way, \eqref{statmatch} implies the functional dependence
\begin{equation}
\thdn = \transf \left( \thup \right)
\end{equation}
The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.

@@

With the experimental setup described and the theory outlined, we now present results comparing the two. Unless stated otherwise, all parameters used in the theory are taken directly from their experimental values listed in Table \ref{paramtable}. In particular, we set $\delup = 0.22$ in all tests, and for the `representative' experiments, we set $\dratdn = 0.24$ and $\epsup = 0.17$.

\subsection{Calibration of the inverse temperature}

All parameters that appear in the theoretical model follow directly from experimental parameters, with the exception of the inverse temperature, $\thup$, of the incoming flow and the parameter $\lamfac$, which represents the number of characteristic wavelengths in the periodic domain. The parameter $\lamfac$ can be chosen in a sensible way by simply selecting an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved. Hence, we will simply chose $\lamfac = \Lambda/2$ throughout, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths. The incoming inverse temperature, $\thup$, however, is more difficult to pin down without using information from the experiments. Our strategy is to use the outgoing skewness as the main diagnostic to determine a realistic range for $\thup$. That is, for an input $\thup$, the downstream inverse temperature $\thdn$ is determined by \eqref{statmatch}, which ultimately sets the skewness of the outgoing wave-field.
